---
title: 'Lab 6 Part 1: Spatial variograms & kriging'
author: "Elmera Azadpour"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap) # for interactive graph
library(gstat) # for kriging
library(stars) #grid for new location predictions
```

## lets read in our spatial data: depth to gw and california counties
```{r}
ca_dgw <- read_sf(here("ca_dgw"), layer = "F2013_DBGS_Points_20150720_093252") %>%  
  clean_names() # measurements are in ft

# check projection
st_crs(ca_dgw) # "EPSG",4326

ca_counties <- read_sf(here("ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name)

# check projections of polygons
st_crs(ca_counties) # "EPSG",3857
# lets use st_transform to transform counties to match crs of dgw


ca_counties <- st_transform(ca_counties, st_crs(ca_dgw)) # transformed

st_crs(ca_counties) # great, now its "EPSG",4326 
```

## Now lets make some graphs & maps
```{r}
ggplot() +
  geom_sf(data = ca_counties) +
  geom_sf(data = ca_dgw, aes(color = dgbs))
```

## lets make interactive map
```{r}
tmap_mode("view")

tm_shape(ca_dgw) + 
  tm_dots("dgbs")

```

## lets isolate to San Joaquin county
```{r}
sj_county <- ca_counties %>% 
  filter(name == "San Joaquin")

## find intersection from dgw and outline of sj county
sj_depth <- ca_dgw %>% 
  st_intersection(sj_county)

plot(sj_depth)
plot(sj_county)


ggplot() +
  geom_sf(data = sj_county) +
  geom_sf(data = sj_depth, aes(color = dgbs))
```

## Check for duplicates, you could remove, take mean value
- first lets find the duplicates! 
```{r}
well_duplicates <- sj_depth %>% 
  get_dupes(latitude, longitude) ## 2 points, pairs of locations where there is duplicates

# we will just remove
sj_depth <- sj_depth %>% 
  filter(!local_well %in% well_duplicates$local_well)

sj_depth %>% 
  get_dupes(latitude, longitude)
```

## lets create a vereogram
- a model that describes the correlation between observation in space basd on how far apart they are
- this gives us information on weight
- then make a new grid
- apply spacial kreiging 
```{r}
## creating variogram
sj_dgw_vgm <- variogram(dgbs ~ 1, data = sj_depth)

plot(sj_dgw_vgm) 
# as distance between observations increases, the increase in semivariance (aka there is a decrease in the correlation between values of those observations) 

sj_dgw_vgm_fit <- fit.variogram(sj_dgw_vgm, model = vgm(nugget = 20, psill = 3000, range = 30, model = "Gau"))

sj_dgw_vgm_fit

plot(sj_dgw_vgm, sj_dgw_vgm_fit) #overall seems to do a good job

```
## lets do some spatial kriging (interpolation)
```{r}
# make a grid that only contains an area that is relvaent over SJ county 
sj_grid <- st_bbox(sj_county) %>% 
  st_as_stars(dx = 0.001, dy = 0.001) %>% 
  st_set_crs(4326) %>% 
  st_crop(sj_county)

plot(sj_grid)
```
## now lets make predicts 
```{r}
sj_dgw_krige <- krige(dgbs ~ 1, sj_depth, sj_grid, model = sj_dgw_vgm_fit)

plot(sj_dgw_krige)
```

